Extended Data

### 2021-10-17 doc/manuscript/Dow_manuscript_Nature.Rmd
minor edits

For ring- and diffuse- porous species at both sites, warmer $T_{max}$ in the critical temperature window resulted in earlier achievement of growth milestones (Figs. `r schematic_summary`, `r EDF_pheno_Tmax_climwin`).
Consistent with findings from previous studies[@parmesan_globally_2003, @ibanez_forecasting_2010,@friedl_tale_2014], the DOY of leaf greenup advanced with increasing $T_{max}$ during the critical temperature window (p=0.001). 
Specifically, greenup DOY advanced 4.5 days/$^\circ$C at SCBI and 2.4 days/ $^\circ$C at Harvard Forest, and mid-greenup also advanced, but to a lesser extent (Fig. `r DOY_timing`).
Similarly, at both sites, $DOY_{25}$, $DOY_{50}$, $DOY_{75}$, and $DOY_{g_{max}}$ all decreased with mean $T_{max}$ during the critical temperature window (Figs. `r schematic_summary`, `r DOY_timing`; `r EDF_pheno_all`).
Specifically, at SCBI, RP growth advanced 1.9, 1.5, and 1.1 days/ $^\circ$C for $DOY_{25}$, $DOY_{50}$, and $DOY_{75}$, respectively, while DP growth advanced 3.5, 3.5, and 3.6 days/ $^\circ$C for $DOY_{25}$, $DOY_{50}$, and $DOY_{75}$, respectively.
At Harvard Forest, RP growth advanced 2.8, 5.0, and 7.2 days/ $^\circ$C for $DOY_{25}$, $DOY_{50}$, and $DOY_{75}$, respectively, while DP growth advanced 7.9, 7.3, and 6.6 days/ $^\circ$C for $DOY_{25}$, $DOY_{50}$, and $DOY_{75}$, respectively.

@@

Together, our results demonstrate that warmer spring temperatures in the temperate deciduous forests of eastern North America advance the phenology of tree stem growth but have little effect on annual woody productivity (Figs. `r schematic_summary`- `r tree_rings`). 
The observed phenological advance in the start of stem growth under warmer springs parallels phenological advances observed for leaf development (Fig. `r DOY_timing`)[@keenan_net_2014;@finzi_carbon_2020] and NEE [@keenan_net_2014; @finzi_carbon_2020].
However, inconsistent with the concept that an earlier start to growth would increase annual woody productivity, we demonstrate that warmer springs advance the end to the period of most active stem growth and have negligible effect on total annual growth for most species and locations (Fig. `r tree_rings`). 
Our observations suggest that the cessation of rapid stem expansion, which occurs mid-summer near the time of peak leaf greenness (`r EDF_growth_curves_all`)[@finzi_carbon_2020], is likely driven by cues other than carbon limitation, such as daylength or sink limitation, which also play an important role in autumn leaf senesence[@zani_increased_2020;@keenan_timing_2015].
Rather than growing season length, the primary effect of temperature on annual growth is a negative effect of $T_{max}$ during the peak growing season, which is associated with of drought stress[@dorangeville_drought_2018].
Warm springs may also augment summer drought stress in some times and places, effectively canceling out any positive effects of an extended growing period[@buermann_widespread_2018]; however, spring temperatures and summer Standardized Precipitation Evapotranspiration Index[@vicente-serrano_multiscalar_2010]) were uncorrelated within our dendrometer band analysis.

@@

Thus, a distinction between interannual variation and directional change may be critical when considering how directional climate change is likely to affect tree growth and ecosystem C dynamics. 
As discussed above, temporal lags between C uptake and release imply that the full effects of warm spring temperatures on forest woody productivity and carbon cycling are unlikely to be apparent in analyses of interannual variation (including this analysis).
Moreover, acclimation of trees to warming temperatures[**REFS**] <!--prevsioulsy we had these, but I'm not sure they really work: [@zhang_drought_2021; @gessler_way_2020]--> and, on longer time scales, species adaptations and shifts in forest composition[**REFS**] are
If we look across spatial gradients where the latter have had time to play out, we see that longer growing seasons are correlated with greater tree growth[@anderson-teixeira_joint_2021], woody productivity[@banburymorgan_global_2021], and NEE [@churkina_spatial_2005].
Thus, warming spring temperatures are expected to increase the biophysical potential for annual tree growth, but that potential is not being realized on an interannual time frame.

>>>>

For ring- and diffuse- porous species at both sites, warmer $T_{max}$ in the critical temperature window resulted in earlier achievement of growth milestones (Figs. `r schematic_summary`, `r EDF_pheno_Tmax_climwin`).
Consistent with findings from previous studies[@parmesan_globally_2003, @ibanez_forecasting_2010,@friedl_tale_2014], the DOY of leaf greenup advanced with increasing $T_{max}$ during the critical temperature window (p= 0.001 at SCBI; p= 0.1 at HF). 
Specifically, greenup DOY advanced 4.5 days/$^\circ$C at SCBI and 2.4 days/ $^\circ$C at Harvard Forest, and mid-greenup also advanced, but to a lesser extent (Fig. `r DOY_timing`).
Similarly, at both sites, $DOY_{25}$, $DOY_{50}$, $DOY_{75}$, and $DOY_{g_{max}}$ all decreased with mean $T_{max}$ during the critical temperature window (Figs. `r schematic_summary`, `r DOY_timing`; `r EDF_pheno_all`).
Specifically, at SCBI, RP growth advanced 1.9, 1.5, and 1.1 days/ $^\circ$C for $DOY_{25}$, $DOY_{50}$, and $DOY_{75}$, respectively, while DP growth advanced 3.5, 3.5, and 3.6 days/ $^\circ$C for $DOY_{25}$, $DOY_{50}$, and $DOY_{75}$, respectively.
At Harvard Forest, RP growth advanced 2.8, 5.0, and 7.2 days/ $^\circ$C for $DOY_{25}$, $DOY_{50}$, and $DOY_{75}$, respectively, while DP growth advanced 7.9, 7.3, and 6.6 days/ $^\circ$C for $DOY_{25}$, $DOY_{50}$, and $DOY_{75}$, respectively. *(This might be better as a table. Add to an existing one? May be able to add to mean_growth_parameters table)*

@@

Together, our results demonstrate that warmer spring temperatures in the temperate deciduous forests of eastern North America advance the phenology of tree stem growth but have little effect on annual woody productivity (Figs. `r schematic_summary`- `r tree_rings`). 
The observed phenological advance in the start of stem growth under warmer springs parallels phenological advances observed for leaf development (Fig. `r DOY_timing`)[@keenan_net_2014;@finzi_carbon_2020] and NEE [@keenan_net_2014; @finzi_carbon_2020].
However, inconsistent with the concept that an earlier start to growth would increase annual woody productivity, we demonstrate that warmer springs advance the end to the period of most active stem growth and have negligible effect on total annual growth for most species and locations (Fig. `r tree_rings`). 
Our observations suggest that the cessation of rapid stem expansion, which occurs mid-summer near the time of peak leaf greenness (`r EDF_growth_curves_all`)[@finzi_carbon_2020], is likely driven by cues other than carbon limitation, such as daylength or sink limitation, which also play an important role in autumn leaf senesence[@zani_increased_2020;@keenan_timing_2015].
Rather than growing season length, the primary effect of temperature on annual growth is a negative effect of $T_{max}$ during the peak growing season, which is associated with drought stress[@dorangeville_drought_2018].
Warm springs may also augment summer drought stress in some times and places, effectively canceling out any positive effects of an extended growing period[@buermann_widespread_2018]; however, spring temperatures and summer Standardized Precipitation Evapotranspiration Index[@vicente-serrano_multiscalar_2010]) were uncorrelated within our dendrometer band analysis.

@@

Thus, a distinction between interannual variation and directional change may be critical when considering how directional climate change is likely to affect tree growth and ecosystem C dynamics. 
As discussed above, temporal lags between C uptake and release imply that the full effects of warm spring temperatures on forest woody productivity and carbon cycling are unlikely to be apparent in analyses of interannual variation (including this analysis).
Moreover, acclimation of trees to warming temperatures[**REFS**] <!--prevsioulsy we had these, but I'm not sure they really work: [@zhang_drought_2021; @gessler_way_2020]--> and, on longer time scales, species adaptations and shifts are likely to alter forest composition[**REFS**].
If we look across spatial gradients where the latter have had time to play out, we see that longer growing seasons are correlated with greater tree growth[@anderson-teixeira_joint_2021], woody productivity[@banburymorgan_global_2021], and NEE [@churkina_spatial_2005].
Thus, warming spring temperatures are expected to increase the biophysical potential for annual tree growth, but that potential is not being realized on an interannual time frame.
